---
title: "Homework 6"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r global_options, include=FALSE}
library(knitr)
library(tidyverse)
library(colorspace)
library(ggforce)
# for convenient piping assignment
library(magrittr)
opts_chunk$set(fig.align="center", fig.height=4, fig.width=5.5)

# data prep:
olympics <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-07-27/olympics.csv')
olympics_2002 <- olympics %>%
  filter(year == 2002, season == "Winter") %>%
  select(sex) %>%
  count(sex) %>%
  pivot_wider(names_from = sex, values_from = n)

```

*Enter your name and EID here*

Connor Hanna
cdh3663

**This homework is due on Mar. 7, 2022 at 11:00am. Please submit as a pdf file on Canvas.**

**Problem 1:  (6 pts)** We will work with the dataset `olympics_2002` that contains the count of each sex for the 2002 Winter Olympics in Salt Lake City and has been derived from the `olympics` dataset. More information about the dataset can be found at https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-07-27/readme.md

```{r}
olympics_2002
```
Follow these steps and display the modified dataframe after each step:

1. Rearrange the dataframe into long form. The resulting dataframe will have two columns, which you should call `sex` and `count`. There will be two rows of data, one for female and one for male athletes.
2. Create a new column with the percent for each `sex`
3. Rename the values in `sex` to "Female" and "Male". 

```{r}
# your code here
olympics_2002 %<>%
  pivot_longer(cols = c("F","M"), names_to = "sex", values_to = "count")
olympics_2002
```
```{r}
# your code here
olympics_2002 %<>%
  mutate(sex_percent = count / sum(count))
olympics_2002
```
```{r}
# your code here
# having fun with methods of assignment
olympics_2002 <-
olympics_2002 |> 
  mutate(
    sex = replace(sex, sex == "F", "Female"),
    sex = replace(sex, sex == "M", "Male")
    )
olympics_2002
```


**Problem 2: (4 pts)** 

Now make a pie chart of the dataset you generated in Problem 1. Use `theme_void()` to remove all distracting elements. 

```{r}
# your code here
olympics_2002 |>
  ggplot(
    aes(
      x0 = 0, y0 = 0,
      r0 = 0, r = 1,
      amount = sex_percent,
      fill = sex
    )
  ) + 
      geom_arc_bar(stat = "pie") +
      theme_void() + 
      coord_fixed() + 
      labs(title = "Sex Ratio in the 2002 Olympics")
```













