---
title: "Homework 1"
output: pdf_document
---

```{r global_options, include=FALSE}
library(knitr)
library(ggplot2)
opts_chunk$set(fig.align="center", fig.height=4, fig.width=5.5)
```

*Enter your name and EID here*

Connor Hanna   
cdh3663

**This homework is due on Jan. 24, 2022 at 11:00am. Please submit as a pdf file on Canvas.**


**Problem 1: (4 pts)** Demonstrate basic command of Markdown by creating a bulleted list with three items, a numbered list with three items, and a sentence that has one word in bold and one word in italics.

*Bulleted List*

* Robert De Niro
* Jean Reno
* Stellan Skarsgaard

*Numbered List*

1. Robert De Niro
2. Jean Reno
3. Stellan Skarsgaard

Sean Bean is usually typecast as *the character who* **dies**. 

**Problem 2: (3 pts)** The `economics` dataset contains various time series data from the US economy:

```{r}
head(economics)
```

Use ggplot to make a line plot of the total population (`pop`, in thousands) versus time (column `date`).

```{r}
ggplot(economics, aes(date, pop)) + 
  geom_line()
```

**Problem 3: (3 pts)** Again using the `economics` dataset, now make a scatter plot (using `geom_point()`) of the number of unemployed versus the total population (`pop`), and color points by date.

```{r}
ggplot(economics, aes(pop, uempmed, color = date)) +
  geom_point()
```
