---
title: "Homework 10"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r global_options, include=FALSE}
library(knitr)
library(tidyverse)
library(ggdendro)
library(sf)
library(colorspace)
opts_chunk$set(fig.align="center", fig.height=4.326, fig.width=7)
```

*Enter your name and EID here*
Connor Hanna
cdh3663

**This homework is due on April 25, 2022 at 11:00am. Please submit as a pdf file on Canvas.**

**Problem 1: (3 pts)** For Problem 1, we will be using `happiness` taken from the World Happiness Report. You can read more about the dataset here: https://www.kaggle.com/datasets/unsdsn/world-happiness.

```{r message = FALSE}
# data preparation
happiness <- read_csv("https://wilkelab.org/SDS375/datasets/happiness.csv")
head(happiness)
```

a) Perform hierarchical clustering of the countries and calculate the distance matrix. You do not need to display the distance matrix.

b) Display clustering results in a dendrogram.

```{r}
# calculating the distance matrix 
dist_out <-
  happiness |> 
    column_to_rownames(var = "country") |>
    scale() |>
    dist(method = "euclidean")

# hierarchical clustering output
hc_out <- hclust(
  dist_out, method = "average"
)

# displaying the results in a dendrogram
ggdendrogram(hc_out, rotate = TRUE)
```


**Problem 2: (3 pts)** Use the clustering results you found in Problem 1 for Problem 2. 

a) Assign clusters by cutting the dendrogram.

b) Plot a scatter plot for two numeric variables of your choice from `happy_data` and add cluster info into scatterplot.

c) Interpret the plot.

```{r, message=FALSE}
# cutting clusters using cutree
cluster <- cutree(hc_out, k = 3)
cluster 

happiness |>
  left_join(
    tibble(
      country = names(cluster),
      cluster = factor(cluster)
    )
  ) |>
  ggplot(aes(GDP, family_score)) +
  geom_point(aes(color = cluster))
```

Looking at the chart, it looks like GDP and family score are closely aligned. Rich countries in the red cluster tend to have high scores in both variables. At the risk of drawing a causal connection from a simple correlation, it would seem that higher productivity enables workers to spend more time at home. 

**Problem 3: (4 pts)** For Problem 3, we will work with the dataset `texas_income`.

a) Bin the `median_income` column into 3 bins (20K-40K, 40K-60K, 60K-90K). Hint: use `case_when()`.

b) Make a choropleth map of Texas counties colored by median income bin (3 total colors).

c) Use an appropriate color scale and use a theme that shows longitude and latitude (nearly any theme other than `theme_void()` will work).

```{r message = FALSE}
# data preparation
texas_income <- readRDS(url("https://wilkelab.org/SDS375/datasets/Texas_income.rds"))
head(texas_income)
```

```{r}
# binning median_income values
texas_income <-
  texas_income |>
    mutate(
      income_bin = case_when(
        median_income < 40000 ~ "low",
        median_income < 60000 ~ "medium",
        median_income < 90000 ~ "high",
        TRUE ~ "NA"
        )
           )

# making a chloropleth map of Texas using the income bins
library(viridis)

ggplot(texas_income, aes(fill = income_bin)) + 
  geom_sf() + 
  scale_fill_manual(values = c("#50A254", "#A9F8AD", "#73C277", "#B8AD4D")) +
  theme_minimal()
```

